---
title: "Airbnb Bookings Analysis"
subtitle: "SVM and Nueral Network Model Performance Analysis"
author: "Pamela Afful"
format:
  revealjs:
    theme: moon
    slide-number: true
    transition: slide
    toc: true
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
---

```{r lib_read, eval=T, echo=F}
#| message: false
#| warning: false
library(dplyr)
library(knitr)
library(plotly)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(zoo)
library(tidyverse)
library(GGally)
library(modelsummary)
library(stargazer)
library(gtsummary)
library(patchwork)
library(glmnet)
library(lmridge)
library(tidyr)
library(caret)
library(dplyr)
library(interactions)
library(kableExtra)
library(patchwork) 
library(knitr)
library(broom)
library(car)
library(maps)
library(RColorBrewer)
library(tree)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ranger)
library(randomForest)
library(gbm)
library(Metrics)
library(lattice)
library(PRROC)
library(yardstick)
library(ROCR)
library(readr)
library(e1071)
```

# Problem Statement

A new competitor in the private residential market is interested in the Airbnb market in Cape Town and would like an analysis on the key predictors of demand of/ availability of Airbnbs. Demand in this case is proxied by the target variable, fully_booked_30, which is a binary variable that populates 1 if the Airbnb is fully booked in the next 30 days and 0 if not. This is done by using and comparing classification models:

-   SVM with a linear kernel

-   SVM with a radial kernel

-   Feed-forward Neural Network

# Exploratory Analysis

The data is highly imbalanced in favour of the negative class. Since the modelling aims to predict availability, this will be treated as the positive class in modelling- this is the target class.

The data used was listings data containing key attributes across Airbnb IDs.

```{r data_read, eval=T, echo=F}
#| message: false
#| warning: false
#| 
listings_data<- read_csv("listings (1).csv")

listings_data_cleaned=listings_data %>% mutate(host_since=lubridate::ymd(host_since),
                         host_response_time=factor(host_response_time,levels=c("within an hour","within a few hours","within a day", "a few days or more")),
                         host_is_superhost=factor(host_is_superhost,levels=c('TRUE','FALSE')),
                         room_type=factor(room_type,levels=unique(listings_data$room_type)),
                         instant_bookable=factor(instant_bookable,level=c('TRUE','FALSE')),
                         fully_booked_30=factor(fully_booked_30, levels=c('yes','no')),
                         host_verifications=factor(case_when(host_verifications %in% c("['phone']") ~'phone_only',
                                                      host_verifications %in% c("['email', 'phone']" ,"['email', 'phone', 'work_email']", "['phone', 'work_email']") ~'phone_email',
                                                      host_verifications %in% c("['email']") ~'email_only'
                           
                         ),levels=c('phone_only','phone_email','email_only'))
                         
                         
                         )



fixed_month_str=ymd("2026-05-01")
fixed_month=floor_date(fixed_month_str,unit='month')

listings_data_cleaned$host_since_months_int =listings_data_cleaned %>% mutate(host_since=ymd(host_since)  ) %>% 
  mutate(host_since_trunc=floor_date(host_since,unit='month')) %>% 
  
  mutate(host_since_months_int= interval(host_since_trunc,fixed_month) %/% months(1) ) %>% 
  
  mutate(current_date= floor_date( today(),unit='month' ) ) %>% 
  
  dplyr::select(host_since_trunc,host_since_months_int,current_date ) %>%
  
  pull(host_since_months_int)

 # create intervals for EDA purposes
a=listings_data_cleaned %>% mutate(host_since_intervals=  cut(host_since_months_int, breaks=c(0,48,96,114,197),labels = c("0-2 yrs", "2-4 yrs", "4-6 yrs",'6+ yrs') )) %>% 
  group_by(host_since_intervals,fully_booked_30) %>% 
  summarize(n_cnt=length(host_id)) %>% 
ggplot(aes(x=fully_booked_30,y=n_cnt, fill=host_since_intervals))+
  geom_col(position="dodge")+
  labs(title = "Fully Booked vs Non Fully Booked Host Age",y='#observations',
       x = "30-day booked status",fill='host age bin')+
  theme_minimal()

a= listings_data_cleaned %>% ggplot(aes(x=price,fill=fully_booked_30)) +
  geom_boxplot()+
    labs(title = "Host Age By 30d Fully Booked Status",x='host age (months)',
       x = "30-day booked status",fill='30 booked status')+
  theme_minimal()
```

```{r target_dist, eval=TRUE, echo=F}
#| message: false
#| warning: false
#| fig-height: 6
#| fig-subcap: Spatial Plot of BNB Listings,Fully Booked 30 Label Distribution
#| fig-width: 8
#| layout-ncol: 2


# Plot the Air BNB locations


cols <- ifelse(listings_data_cleaned$fully_booked_30 == 'no', "red", "blue")


plot(listings_data_cleaned$longitude,
      listings_data_cleaned$latitude,
     xlab = "Longitude",
     ylab = "Latitude",
       pch = 16,
       cex = 0.5,
       col = cols)

legend("topright",
       legend = c("Fully Booked", "Not Fully Booked"),
       col = c("blue", "red"),
       pch = 16,
       bty = "n",
       title = "Fully Booked 30",
       cex = 0.9)



listings_data_cleaned %>% 
  count(fully_booked_30) %>% 
  mutate(
    percentage = n / sum(n),
    label = paste0(round(percentage * 100, 1), "%")
  ) %>%
  ggplot(aes(x = fully_booked_30, y = n)) +
  geom_col(fill = 'steelblue') +
  geom_text(aes(label = label), 
            vjust = -0.5, size = 3) +
  labs(
    title = "Fully Booked vs Non Fully Booked",
    y = '#observations',
    x = "30-day booked status"
  ) +
  theme_minimal()
```

## Feature Distributions

```{r eda_plots, eval=TRUE, echo=FALSE}
#| message: false
#| warning: false
#| fig-height: 8.8
#| fig-width: 15


num_vars=listings_data_cleaned %>% group_by(fully_booked_30) %>% summarize(
  avg_host_response_rate=mean(host_response_rate),
  avg_host_acceptance_rate=mean(host_acceptance_rate),
  avg_host_listings_count=mean(host_listings_count),
  avg_price=mean(price),
  avg_minimum_nights=mean(minimum_nights),
  avg_review_scores_rating=mean(review_scores_rating)
)

host_age_dist=listings_data_cleaned %>% ggplot(aes(x=host_since_months_int,fill=fully_booked_30))+
  geom_boxplot()+
  labs(title='Host Age Distribution by 30d Fully Booked Status',x='Host Age (months)',fill='Status')+theme_minimal()
  


vars <- c(
  "host_response_rate",
  "host_acceptance_rate",
  "host_listings_count",
  "price",
  "minimum_nights",
  "review_scores_rating"
)



# Initialize empty list
plot_list <- list()

# Loop through variables
for (var in vars) {
  
  # Create summary dataframe dynamically
  df_plot <- listings_data_cleaned %>%
    group_by(fully_booked_30) %>%
    summarise(mean_value = mean(.data[[var]], na.rm = TRUE))
  
  # Create plot
  p <- ggplot(df_plot, aes(x = fully_booked_30, y = mean_value)) +
    geom_col(fill = "steelblue") +
    labs(
      title = paste("Average", gsub("_", " ", var), "by Booking Status"),
      x = "30-day booked status",
      y = paste("Average", gsub("_", " ", var))
    ) +
    theme_minimal()
  
  # Store in list
  plot_list[[var]] <- p
}

# View one example

vars <- c(
  "host_response_rate",
  "host_acceptance_rate",
  "host_listings_count",
  "price",
  "minimum_nights",
  "review_scores_rating"
)

(plot_list[["price"]]+plot_list[["minimum_nights"]]+plot_list[["review_scores_rating"]])/ (plot_list[["host_acceptance_rate"]]+plot_list[["host_response_rate"]]+plot_list[["host_listings_count"]])+host_age_dist
```

# Model Training

## Linear Kernel SVM

```{r, svm_model_perf, eval=T, echo=F}
liner_svm_model=readRDS("svm_model.rds")
liner_svm_model$performances
```
